### 2016-05-27 2016_ICML/poster/defines/boxes.tex

%set boxes

\setbeamerfont{block title}{shape=\bfseries,size=\large,parent={structure,block body}}

\newcommand{\drawboxes}{
	%boxes
	\setbeamercolor*{block body}{fg=black,bg=white}
	\setbeamercolor*{block title}{fg=black,bg=white}

	\setbeamertemplate{block begin}{
	\vskip.75ex
	\begin{beamercolorbox}[rounded=true,shadow=true,leftskip=1cm,colsep*=.75ex]{block title}%
		\usebeamerfont*{block title}\insertblocktitle
	\end{beamercolorbox}%
	{\ifbeamercolorempty[bg]{block body}{}{\nointerlineskip\vskip-0.5pt}}%
	\usebeamerfont{block body}%
	\begin{beamercolorbox}[rounded=true,shadow=true,colsep*=.75ex,sep=.75ex,vmode]{block body}%
		\ifbeamercolorempty[bg]{block body}{\vskip-.25ex}{\vskip-.75ex}\vbox{}%
	}
			\setbeamertemplate{block end}{
	\end{beamercolorbox}
			\vspace*{2ex}
	}
}