### 2016-05-27 2016_ICML/poster/defines/header.tex

\pgfdeclareimage[height=4.8cm]{dfki-logo}{./defines/DFKI_Schrift}

\setbeamertemplate{headline}
{%
  \setbeamercolor{mycolor}{fg=black,bg=white}
   \begin{beamercolorbox}{mycolor}
    \centering{\parbox{70cm}{\centering{\huge{\textbf{\inserttitle}}}}}\\[5mm]
    \centering{\parbox{70cm}{\centering{\Large{\insertauthor\\[.5cm] University Bremen,
Robotics Research Group, Robert-Hooke-Str.~1, 28359 Bremen,
Germany\hspace{2cm}}}}}

  \end{beamercolorbox}%
  \vspace{-6cm}
}