### 2016-05-31 2016_ICML/poster/defines/header.tex
Working on poster layout

\pgfdeclareimage[height=2.8cm]{dfki-logo}{./defines/Bosch-Logo}

\setbeamertemplate{headline}
{%
  \setbeamercolor{mycolor}{fg=black,bg=white}
   \begin{beamercolorbox}{mycolor}
    \centering{\parbox{70cm}{\centering{\huge{\textbf{\inserttitle}}}}}\\[5mm]
    \centering{\parbox{70cm}{\centering{\Large{\insertauthor\\[.5cm] University Bremen,
Robotics Research Group, Robert-Hooke-Str.~1, 28359 Bremen, Germany\\
Robert Bosch GmbH, Corporate Research CR/AEY2, 70442 Stuttgart, Germany
}}}}

  \end{beamercolorbox}%
  \vspace{-6cm}
}